---
title: "BIOMASS"
output: 
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The package

R package for estimating above-ground biomass and its uncertainty in tropical forests. 

Contains functions to estimate aboveground biomass/carbon and its uncertainty in tropical forests. 
These functions allow to :  
1. retrive and correct the taxonomy,  
2. estimate the wood density and its incertainty,  
3. construct height-diameter models,  
4. estimate the above-ground biomass/carbon at the stand level with asociated uncertainty.



For more informations, this is the [article](https://besjournals.onlinelibrary.wiley.com/doi/pdf/10.1111/2041-210X.12753) talking about this package.

There is the [vignette](https://CRAN.R-project.org/package=BIOMASS/vignettes/VignetteBiomass.html) of the package, and the [reference manual](https://CRAN.R-project.org/package=BIOMASS/BIOMASS.pdf).


## Citation

To cite BIOMASS, please use citation("BIOMASS").


## Install BIOMASS

The latest released version from CRAN with

```{r, eval=FALSE}
install.packages("BIOMASS")
```

To use it :
```{r, eval=FALSE}
library("BIOMASS")
```
