---
title: "Manage tree and plot coordinate with BIOMASS"
author: "Arthur Pere"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty: 
    number_sections: yes
    toc: yes
    highlight: vignette
    self_contained: yes
    theme: cayman
vignette: >
  %\VignetteIndexEntry{Manage tree and plot coordinate with BIOMASS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo = TRUE,
  comment = "#>", fig.align = "center"
)
require(BIOMASS)
require(knitr)
```


# Requirement

You need the following information to use the following function in your analysis :
  
  - Dimension of the plots
  - Coordinate GPS of minimum 3 corners with the corresponding relative coordinates
  - The origin corner
  - The X, Y directions

```{r}
trees <- read.csv(system.file("external", "NouraguesPlot.csv",
  package = "BIOMASS", mustWork = T
))
```

```{r echo=FALSE}
kable(head(trees), digits = 3, row.names = F, caption = "Head of the table trees")
```






```{r, fig.cap="Plot the coordinate"}
coord <- read.csv(system.file("external", "Coord.csv",
  package = "BIOMASS", mustWork = T
))

plot(coord[, c("Long", "Lat")], asp = 1)
```
```{r echo=FALSE}
kable(head(coord), digits = 3, row.names = F, caption = "Head of the table coord")
```





# To manage the plots
## Import the dataset and visualisation

You can see that the corner coordinates are spread, and we can correct this.

The plot is referenced in the longitude latitude coordinate so you must have the package `proj4` if you are in this situation. If you have projected coordinate, you can continue with the `projCoord` argument instead of `longlat` argument. 

## Correct the GPS coordinate
```{r, cache=FALSE}
correct_plot <- correctCoordGPS(
  longlat = coord[, c("Long", "Lat")],
  coordRel = coord[, c("xRel", "yRel")],
  rangeX = c(0, 100), rangeY = c(0, 100), drawPlot = T,
  maxDist = 10, rmOutliers = T
)

str(correct_plot, max.level = 1)
```


The output of the function is a list with a data.frame `corner` it's the corner of the plot, `polygon` the spatial polygon and `outliers` the vector with the line number of the outliers.

The outliers are calculated by a measure of distance between the predicted points and the GPS points. If this distance is higher than the value of `maxDist`, the point is considered like outliers.


## Numbering the corner
We have to number the corner of the plot, it is working if we have exactly 4 points for each plot, so we have to do the correctCoordGPS before if we have not the correct number of points.

```{r}
coord_num <- numberCorner(
  projCoord = correct_plot$cornerCoords,
  plot = rep("NB1", 4),
  origin = c(F, F, F, T),
  clockWise = T
)


plot(coord_num[, c("X", "Y")], asp = 1)
text(coord_num[, c("X", "Y")], labels = coord_num[, "corner"], pos = 2, offset = 0.2)
```

On the graph, you can noted than the corner number 1 the origin of the plot.


## Cut the plot in multiple subplot

```{r}
subplot <- cutPlot(
  projCoord = coord_num[, c("X", "Y")],
  plot = coord_num[, c("plot")],
  corner = coord_num[, c("corner")],
  gridsize = 25, dimX = 100, dimY = 100
)
```

```{r echo=FALSE}
kable(head(subplot))
```


# Trees managements
## Attribute the trees to the subplot
attribute the trees to the to the subplot

```{r}
trees$subplot <- attributeTree(trees[, c("xRel", "yRel")], rep("NB1", nrow(trees)), subplot)
```



## Calculate the AGB and spatialisation
```{r}
trees$AGB <- computeAGB(trees$D, trees$WD, H = trees$H)

AGB <- summaryByPlot(trees$AGB, trees$subplot, drawPlot = T, subplot = subplot)

print(AGB)
```











