---
title: "Predict maps of AGBD based on inventory and LiDAR data"
author: "Dominique Lamonica"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty: 
    number_sections: yes
    toc: yes
    highlight: vignette
    self_contained: yes
vignette: >
  %\VignetteIndexEntry{Predict maps of AGBD based on inventory and Lidar data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo = TRUE,
  comment = "#>", fig.align = "center"
)
require(BIOMASS)
require(knitr)
require(ggplot2)
require(dplyr)
```

# Overview

- what data + updated workflow
- bayesian method to propagate uncertainties, using brms (ref), parallel using doFuture (ref)
- model, ref Hunka
- example with 4 plots to illustrate, but keep in mind it is very likely not enough data to get proper and reliable final predictions !

# Previous steps (see other vignettes for details)

## Load inventory, plot coordinates and LiDAR data
```{r}
data("NouraguesTrees")
data("NouraguesCoords")
nouraguesRaster <- terra::rast(system.file("extdata", "NouraguesRaster.tif",package = "BIOMASS", mustWork = TRUE))
```

## Compute AGBD
```{r, eval=FALSE}
# Height
data("NouraguesHD")
brm_model <- modelHD(
  D = NouraguesHD$D, H = NouraguesHD$H,
  method = "log2",
  bayesian = TRUE, useCache = TRUE)

# Wood density
Taxo <- readRDS(file = "saved_data/Taxo_vignette.rds")
NouraguesTrees$GenusCorrected <- Taxo$genusAccepted
NouraguesTrees$SpeciesCorrected <- Taxo$speciesAccepted
NouraguesTrees$family <- Taxo$familyAccepted
wood_densities <- getWoodDensity(
  genus = NouraguesTrees$GenusCorrected,
  species = NouraguesTrees$SpeciesCorrected,
  family = NouraguesTrees$family,
  stand = NouraguesTrees$Plot
)
NouraguesTrees$WD <- wood_densities$meanWD

error_prop_4plots <- AGBmonteCarlo(
  D = NouraguesTrees$D, WD = NouraguesTrees$WD,
  HDmodel = brm_model,
  Dpropag = "chave2004",
  errWD = wood_densities$sdWD)
# keep only 50 iterations per tree for vignette example
error_prop_4plots$AGB_simu <- error_prop_4plots$AGB_simu[,1:50]
```
```{r load_error_prop, include=FALSE}
error_prop4plots <- readRDS(file = "saved_data/error_prop4plots.rds")
```

## Spatialize AGBD

```{r}
# divide plots into subplots
multiple_subplots <- divide_plot(
  grid_size = 25, 
  corner_data = NouraguesCoords,
  rel_coord = c("Xfield","Yfield"), proj_coord = c("Xutm","Yutm"), corner_plot_ID = "Plot",
  tree_data = NouraguesTrees, tree_coords = c("Xfield","Yfield"), tree_plot_ID = "Plot"
)

# check with raster, optional
multiple_checks <- check_plot_coord(
  corner_data = NouraguesCoords, # NouraguesCoords contains 4 plots
  proj_coord = c("Xutm", "Yutm"), rel_coord = c("Xfield", "Yfield"),
  trust_GPS_corners = TRUE,
  plot_ID = "Plot",
  tree_data = NouraguesTrees, tree_coords = c("Xfield","Yfield"),
  prop_tree = "D", tree_plot_ID = "Plot",
  ref_raster = nouraguesRaster)

# compute AGBD estimates and their uncertainty per subplot
subplot_AGBD <- subplot_summary(
  subplots = multiple_subplots,
  AGB_simu = error_prop4plots$AGB_simu, draw = F
)
```

## Spatialize LiDAR metric
```{r}
# quick plot to visualise plot corners in the landscape
terra::plot(nouraguesRaster)
points(NouraguesCoords$Xutm, NouraguesCoords$Yutm, col ="red", pch = 20)

# get CHM median values for each suplot
raster_summary <- subplot_summary(
  subplots = multiple_subplots,
  ref_raster = nouraguesRaster, raster_fun = median, na.rm = T)
chm_subplot <- raster_summary$tree_summary |>
  rename(raster_metric = z2012_median)
```

# Calibrate model
Gather data for agbd-chm model calibration (ie join predictor-predicted):
```{r}
agbd_subplot <- subplot_AGBD$long_AGB_simu
data_for_calibration <- agbd_subplot %>%
  left_join(chm_subplot, by = "subplot_ID") %>%
  arrange(subplot_ID)
```

then run calibration function, here parallelized on four CPUs:
```{r, eval=FALSE}
model_cal <- calibrate_model(long_AGB_simu = data_for_calibration, nb_rep = 50, useCache = T,
                            plot_model = TRUE, chains = 4, thin = 20, iter = 2500,
                            warmup = 500, cores = 4)
```
```{r load_brms_output, include=FALSE}
model_cal <- readRDS(file = "saved_data/nouragues_model_cal.rds")
```

Let's check inference results:
```{r}
summary(model_cal)
```
(maybe ppc ?)

# Predict AGBD map on the LiDAR footprint
```{r}
map_agbd <- predict_map(fit_brms = model_cal,
                         pred_raster = nouraguesRaster,
                         grid_size = 25,
                         raster_fun = median,
                         n_cores = 4,
                         n_post_draws = 100,
                         alignment_raster = NULL,
                         plot_maps = T)
```
If you want to predict on another footprint, you can supply the raster footprint with `alignment_raster` argument.



