---
title: "Spatialized trees and forest stands metrics with BIOMASS"
author: "Arthur Bailly"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty: 
    number_sections: yes
    toc: yes
    highlight: vignette
    self_contained: yes
    theme: cayman
vignette: >
  %\VignetteIndexEntry{Spatialized trees and forest stands metrics with BIOMASS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, echo = TRUE,
  comment = "#>", fig.align = "center"
)
#require(BIOMASS)
devtools::load_all()
require(knitr)
require(ggplot2)
require(data.table)
require(tidyverse)
require(dplyr)
```

/!\\ VIGNETTE IN PROGRESS /!\\

# Overview

BIOMASS enables users to manage their plots by :

* calculating the projected/geographic coordinates of the plot's corners and trees from the relative coordinates (or local coordinates, ie, those of the field) 

* visualizing the plots

* validating plot's corner and tree coordinates with LiDAR data

* dividing plots into subplots

* summarizing the desired information at subplot level


# Required data

Two data frames are required for the rest : 

* a **data frame of corner coordinates**, containing at least :
  - the name of the plots if there are several plots
  - the coordinates of the plot's corners in the geographic or projected coordinate system (the GPS coordinates or another projected coordinates)
  - the coordinates of the plot's corners in the relative coordinate system (the local or field coordinates)

In this vignette, we will use the Nouragues dataset as an exemple :

```{r}
corner_coord <- read.csv(system.file("external", "Coord.csv",package = "BIOMASS", mustWork = TRUE))

kable(head(corner_coord), digits = 5, row.names = FALSE, caption = "Head of the table corner_coord")
```

Note that both Lat/Long and xProj/yProj coordinates are included in this dataset but only one of these coordinate systems is needed.  

* a **data frame of trees coordinates**, containing at least :
  - the name of the plots if there are several plots
  - the tree coordinates in the relative coordinate system (the local/field one)
  - the desired information about trees, such as diameter, wood density, height, etc. (see BIOMASS vignette)


```{r}
trees <- read.csv(system.file("external", "NouraguesPlot.csv",package = "BIOMASS", mustWork = TRUE))

kable(head(trees), digits = 3, row.names = FALSE, caption = "Head of the table trees")
```

# Checking plot's coordinates

The user is faced with two situations : 

* The **GPS coordinates** of the plot corners are considered **very accurate** or **enough measurements have been made** to be confident in the accuracy of their mean. In this case, the shape of the plot measured on the field will follow the GPS coordinates of the plot corners when projected into the projected/geographic coordinate system. See [3.1.1](#trustGSPcorner)

* **Too few measurements of the GPS coordinates** of plot corners have been taken and/or are **not reliable**. In this case, **the shape of the plot measured on the field is considered to be accurate** and the GPS corner coordinates will be recalculated to fit the shape and dimensions of the plot (the projected coordinates fit the relative coordinates).  See [3.1.2](#trustPlotShape)

In both case, the use of the `check_plot_coord()` function is recommended as a first step.

## Checking the corners of the plot

The `check_plot_coord()` function handles both situations using the `trust_GPS_corners` argument (= TRUE or FALSE).

You can give either GPS coordinates (with the longlat argument) or another projected coordinates (with the proj_coord argument) for the corner coordinates. 

### If we rely on the GPS coordinates of the corners {#trustGSPcorner}

If **enough coordinates have been recorded for each corner** (for more information, see the [CEOS good practices protocol, section A.1.3.1](https://lpvs.gsfc.nasa.gov/PDF/CEOS_WGCV_LPV_Biomass_Protocol_2021_V1.0.pdf) ), you will need to provide the corner names via the `corner_ID` argument. In this case, each coordinates will be averaging by corner, resulting in 4 reference coordinates. The function can also detect and remove GPS outliers using the rm_outliers and max_dist arguments. 

If only **4 GPS measurements** have been taken **with a high degree of accuracy** (by a geometer, for example), or if you already have averaged your measurements by yourself, you can supply these 4 GPS coordinates to the function.

```{r dpi=200}
check_plot_trust_GPS <- check_plot_coord(
  longlat = corner_coord[c("Long", "Lat")],
  # Or, if exists : proj_coord = corner_coord[c("xProj", "yProj")], 
  rel_coord = corner_coord[c("xRel", "yRel")],
  trust_GPS_corners = T, corner_ID = corner_coord$Corners,
  draw_plot = TRUE,
  max_dist = 10, rm_outliers = TRUE )
```

The two blue arrows represent the relative coordinate system when projected into the projected coordinate system. 


### If we rely on the shape of the plot measured on the field {#trustPlotShape}

Let's degrade the data to simulate the fact that we only have 8 GPS coordinates that we don't trust.

```{r dpi=200}
degraded_corner_coord <- corner_coord[c(1:2,11:12,21:22,31:32),]
  
check_plot_trust_field <- check_plot_coord(
  longlat = degraded_corner_coord[, c("Long", "Lat")],
  # OR proj_coord = corner_coord[c("xProj", "yProj")], 
  rel_coord = degraded_corner_coord[, c("xRel", "yRel")],
  trust_GPS_corners = F,
  draw_plot = TRUE )
```

We can see that the corners of the plot do not correspond to the GPS measurements. In fact, they correspond to the best compromise between the shape and dimensions of the plot and the GPS measurements.

## Tree coordinates in the projected/geographic coordinate system {#tree_coordinates}

Tree coordinates are almost always measured in the relative (local/field) coordinate system. To retrieve them in the projected system, you can supply their relative coordinates using the `tree_df` and `tree_coords` arguments.

```{r}
check_plot_trust_GPS <- check_plot_coord(
  longlat = corner_coord[c("Long", "Lat")],
  # OR proj_coord = corner_coord[c("xProj", "yProj")], 
  rel_coord = corner_coord[c("xRel", "yRel")],
  trust_GPS_corners = T, corner_ID = corner_coord$Corners,
  draw_plot = TRUE,
  max_dist = 10, rm_outliers = TRUE,
  tree_df = trees, tree_coords = c("xRel","yRel") )

check_plot_trust_field <- check_plot_coord(
  longlat = degraded_corner_coord[, c("Long", "Lat")],
  # OR proj_coord = corner_coord[c("xProj", "yProj")], 
  rel_coord = degraded_corner_coord[, c("xRel", "yRel")],
  trust_GPS_corners = F,
  draw_plot = TRUE,
  tree_df = trees, tree_coords = c("xRel","yRel") )
```

Note the difference in corner and tree positions between the two situations.

The tree coordinates can be obtained via the `$tree_proj_coord` output.

```{r}
trees[c("x_proj","y_proj")] <- check_plot_trust_GPS$tree_proj_coord
```

You can also grep and modify the plot via the `$plot_design` output which is a ggplot object. For example, to change the plot title : 

```{r}
plot_to_change <- check_plot_trust_GPS$plot_design
plot_to_change <- plot_to_change + ggtitle("A custom title")
plot_to_change
```

If you want to retrieve the GPS coordinates of the trees in a longitude/latitude format, use this code:

```{r}
tree_GPS_coord <- as.data.frame( proj4::project(check_plot_trust_GPS$tree_proj_coord, proj = check_plot_trust_GPS$UTM_code, inverse = TRUE) )
```


## Integrating LiDAR data

WORK IN PROGRESS

# Dividing plots

Dividing plots into several rectangular sub-plots is performed using the `divide_plot()` function. This function takes the relative coordinates of the 4 corners of the plot to divide it into a grid (be aware that **the plot must be rectangular in the relative coordinates system**, ie, have 4 right angles) :

```{r divide_plot}
subplots <- divide_plot(
  rel_coord = check_plot_trust_GPS$corner_coord[c("x_rel","y_rel")],
  proj_coord = check_plot_trust_GPS$corner_coord[c("x_proj","y_proj")],
  grid_size = 25 # or c(25,25)
  )

kable(head(subplots, 10), digits = 1, row.names = FALSE, caption = "Head of the divide_plot() returns")
```

If you don't have any projected coordinates, just let `proj_coord` = NULL.

For the purposes of summarizing and representing subplots (coming in the next section), the function returns the coordinates of subplot corners but can also **assign to each tree its subplot** with the **`tree_df` and `tree_coords` arguments** :

```{r divide_plot_trees}
subplots <- divide_plot(
  rel_coord = check_plot_trust_GPS$corner_coord[c("x_rel","y_rel")],
  proj_coord = check_plot_trust_GPS$corner_coord[c("x_proj","y_proj")],
  grid_size = 25, # or c(50,50)
  tree_df = trees, tree_coords = c("xRel","yRel")
  )
```

The function now returns a list containing : 

* `sub_corner_coord` : coordinates of subplot corners as previously

* `tree_df` : the tree data-frame with the **subplot_id added in last column**

```{r}
kable(head(subplots$tree_df), digits = 1, row.names = FALSE, caption = "Head of the divide_plot()$tree_df returns")
```

The function can handle as many plots as you wish, using the `corner_plot_ID` and `tree_plot_ID` arguments :

```{r divide_multiple_plots}
multiple_plots <- rbind(check_plot_trust_GPS$corner_coord[c("x_rel","y_rel","x_proj","y_proj")], check_plot_trust_GPS$corner_coord[c("x_rel","y_rel","x_proj","y_proj")])
multiple_plots$plot_ID = rep(c("NB1","NB2"),e=4)
multiple_trees <- rbind(trees,trees)
multiple_trees$plot = rep(c("NB1","NB2"),e=nrow(trees))

multiple_subplots <- divide_plot(
  rel_coord = multiple_plots[c("x_rel","y_rel")],
  proj_coord = multiple_plots[c("x_proj","y_proj")],
  grid_size = 50,
  corner_plot_ID = multiple_plots$plot_ID,
  tree_coord = multiple_trees,
  tree_plot_ID = multiple_trees$plot
)
```

# Summarizing tree metrics at subplot level

Once you've applied the `divide_plot()` function with a valid `tree_coord` argument, you can summarize any tree metric present in divide_plot()$tree_coord at the subplot level with the subplot_summary() function.

```{r subplot_summary}
subplot_metric <- subplot_summary(subplots = subplots,
                                  value = "H") # H is the tree metric

subplot_metric$tree_summary
subplot_metric$polygon
subplot_metric$plot_design
```

By default, the function sums the metric per subplot but you can request any valid function with the `function_name` argument :

```{r subplot_summary_sd}
subplot_metric <- subplot_summary(subplots = subplots,
                                  value = "H", #H is the tree metric
                                  function_name = "median") 
```

The output of the function is a list containing : 

* `$tree_summary` : a summary of the metric per subplot 

* `$polygon` : same as `$tree_summary` with sf polygons in last column 

* `$plot_design` : (if draw_plot=T) a ggplot object that can easily be modified


** DEUX OPTIONS POUR LA SUITE **

* Soit l'utilisateur peut set display_trees = TRUE, et les arbres seront affichés mais dans ce cas, il faut que les colonnes x_proj et y_proj soit bien présentes dans son tree data-frame 

* Soit on laisse l'utilisateur afficher lui-même les arbres car c'est plutôt facile : 

```{r}
custom_plot <- subplot_metric$plot_design
custom_plot + 
  geom_point(data=trees, mapping = aes(x = x_proj, y = y_proj, size = D), shape=1) + 
  ggtitle("The title you want") 
```


```{r subplot_summary_display_trees}
multiple_subplot_metric <- subplot_summary(
  subplots = multiple_subplots,
  value = "H", display_trees = T)


```
